A reverse double negation translation

### 2022-04-07 Notes/A reverse double negation translation/RDNT.tex
Full Proof Lemma 2.7

@@

@@

	A sequent $A\Rightarrow B$ consists of two multi-sets of formulas $A, B$. We will be using the following rules for our classical calculus:\\
	\begin{center}
	\begin{tabular}{lll}
		\AxiomC{\hphantom{x}}
		\RightLabel{Ax}
		\UnaryInfC{$A,\Gamma\Rightarrow \Delta, A$}
		\DisplayProof&
		\AxiomC{\hphantom{x}}
		\RightLabel{L$\bot$}
		\UnaryInfC{$\bot,\Gamma\Rightarrow\Delta$}
		\DisplayProof&
		\\&&\\
		\AxiomC{$A, B,\Gamma\Rightarrow\Delta$}
		\RightLabel{L$\wedge$}
		\UnaryInfC{$A\wedge B, \Gamma\Rightarrow \Delta$}
		\DisplayProof&
		\AxiomC{$\Gamma\Rightarrow\Delta, A$}
		\AxiomC{$\Gamma\Rightarrow\Delta, B$}
		\RightLabel{R$\wedge$}
		\BinaryInfC{$\Gamma\Rightarrow \Delta, A\wedge B$}
		\DisplayProof&
		\\&&\\
		\AxiomC{$A, \Gamma\Rightarrow\Delta$}
		\AxiomC{$B, \Gamma\Rightarrow\Delta$}
		\RightLabel{L$\vee$}
		\BinaryInfC{$A\vee B, \Gamma\Rightarrow \Delta$}
		\DisplayProof&
		\AxiomC{$\Gamma\Rightarrow\Delta, A, B$}
		\RightLabel{R$\vee$}
		\UnaryInfC{$\Gamma\Rightarrow \Delta, A\vee B$}
		\DisplayProof&
		\\&&\\
		\AxiomC{$A\to B, \Gamma\Rightarrow\Delta, A$}
		\AxiomC{$B, \Gamma\Rightarrow\Delta$}
		\RightLabel{L$\to$}
		\BinaryInfC{$A\to B, \Gamma\Rightarrow \Delta$}
		\DisplayProof&
		\AxiomC{$A,\Gamma\Rightarrow\Delta, B$}
		\RightLabel{R$\to$}
		\UnaryInfC{$\Gamma\Rightarrow \Delta, A\to B$}
		\DisplayProof&
		\\&&\\
	\end{tabular}
	\end{center}

@@

	\begin{definition}
		We say that two sequents $S, S'$ are \textit{equi-provable} if $S$ is provable iff $S'$ is provable in intuitionistic logic.
	\end{definition}

	Clearly equi-provability is an equivalence relation.

	\begin{example}
		The sequents $\Rightarrow A\vee B$ and $A\to C, B\to C\Rightarrow C$ are equi-provable. Note that they are not equivalent in the typical sense however, due to the absence of $C$ in the former. 
	\end{example}

@@

		However there is clearly some collection between the two notions:

		\begin{lemma}
			Let $A$ and $B$ be equivalent formulas, i.e. $A\Rightarrow B$ and $B\Rightarrow A$ are provable, then for any $\Gamma, \Delta$ the sequents $A,\Gamma\Rightarrow\Delta$ and $B, \Gamma\Rightarrow\Delta$ as well as $\Gamma\Rightarrow\Delta, A$ and $\Gamma\Rightarrow\Delta, B$ are equi-provable.
		\end{lemma}

@@

	\begin{lemma}
		Let $N$ be a atomic proposition that does not occur in $\Gamma$ or $\Delta$, $A, B$ arbitrary formulas that do not contain $N$, then:
		\begin{enumerate}
			\item $\Gamma\Rightarrow\Delta$ and $\Gamma\Rightarrow\Delta, N$ are equi-provable.
			\item $\Gamma\Rightarrow\Delta, A$ and $\Gamma, A\to N\Rightarrow \Delta, N$ are equi-provable.
		\end{enumerate} 
	\end{lemma}

@@

		2. If $\Gamma\Rightarrow\Delta, A$ is provable we obtain $\Gamma, A\to N\Rightarrow \Delta, N$ through a simple application of cut:
		\begin{center}
			\AxiomC{$\vdots$}
			\noLine
			\UnaryInfC{$\Gamma\Rightarrow\Delta, A$}
			\AxiomC{\hphantom{X}}
			\RightLabel{Ax}
			\UnaryInfC{$A, A\to N\Rightarrow P, A$}
			\AxiomC{\hphantom{X}}
			\RightLabel{Ax}
			\UnaryInfC{$N, A\Rightarrow N$}
			\RightLabel{L$\to$}
			\BinaryInfC{$A, A\to N\Rightarrow N$}
			\RightLabel{Cut}
			\BinaryInfC{$\Gamma, A\to N\Rightarrow\Delta, N$}
			\DisplayProof
		\end{center}
		Suppose on the other hand we have a proof of $\Gamma, A\to N\Rightarrow \Delta, N$. Consider the proof obtained by replacing every occurrence of $N$ with $A$. By induction on the proof height it can be checked that this constitutes a valid proof of $\Gamma, A\to A\Rightarrow\Delta, N$. We obtain $\Gamma\Rightarrow\Delta, A$ through an application fo Cut:
		\begin{center}
			\AxiomC{\hphantom{X}}
			\RightLabel{Ax}
			\UnaryInfC{$A, \Rightarrow A$}
			\RightLabel{R$\to$}
			\UnaryInfC{$\Rightarrow A\to A$}
			\AxiomC{$\vdots$}
			\noLine
			\UnaryInfC{$\Gamma, A\to A\Rightarrow\Delta, A$}
			\RightLabel{Cut}
			\BinaryInfC{$\Gamma\Rightarrow\Delta, A$}
			\DisplayProof
		\end{center}
	\end{proof}

@@

	The first result is that for every sequent there is some equi-provable implicational sequent. This transformation is not strictly necessary for what follows, but it makes the arguments and procedure considerably simpler. It is reminiscent of the definitions of connectors in second order logic.

	Consider the following rules, where $X$ is not atomic, $N$ is a new atom:
	\begin{center}
		\begin{tabular}{lcl}
			$\Delta\Rightarrow \Gamma, A, B$&$\rightsquigarrow$&$\Delta\Rightarrow\Gamma, A\vee B$\\
			$\Delta\Rightarrow A\vee B$&$\rightsquigarrow$&$\Delta, (A\vee B)\to N\Rightarrow N$\\
			$\Delta\Rightarrow A\wedge B$&$\rightsquigarrow$&$\Delta,(A\wedge B)\to N\Rightarrow N$\\
			$\Delta\Rightarrow A\to B$&$\rightsquigarrow$&$\Delta, A\Rightarrow B$\\

			$\Delta, A\wedge B\Rightarrow C$&$\rightsquigarrow$&$\Delta, A, B\Rightarrow C$\\
			$\Delta, A\vee B\Rightarrow C$&$\rightsquigarrow$&$\Delta, (A\to C)\to (B\to C)\to C\Rightarrow  C$\\
			$\Delta, (A\wedge B)\to C\Rightarrow D$&$\rightsquigarrow$&$\Delta, A\to (B\to C)\Rightarrow D$\\
			$\Delta, (A\vee B)\to C\Rightarrow D$&$\rightsquigarrow$&$\Delta, A\to C, B\to C\Rightarrow D$\\
			$\Delta, A\to (B\vee C)\Rightarrow D$&$\rightsquigarrow$&$\Delta, A\to (B\to D)\to (C\to D)\to D\Rightarrow D$\\
			$\Delta, A\to (B\wedge C)\Rightarrow D$&$\rightsquigarrow$&$\Delta, A\to B, A\to C\Rightarrow D$\\
			$\Delta, X\to (A\to B)\Rightarrow C$&$\rightsquigarrow$&$\Delta, X\to N, N\to (A\to B)\Rightarrow C$\\
			$\Delta, A\to (X\to B)\Rightarrow C$&$\rightsquigarrow$&$\Delta, X\to N, A\to (N\to B)\Rightarrow C$\\
			$\Delta, A\to (B\to X)\Rightarrow C$&$\rightsquigarrow$&$\Delta, N\to X, A\to (B\to N)\Rightarrow C$\\
			$\Delta, (X\to A)\to B\Rightarrow C$&$\rightsquigarrow$&$\Delta, N\to X, (N\to A)\to B\Rightarrow C$\\
			$\Delta, (A\to X)\to B\Rightarrow C$&$\rightsquigarrow$&$\Delta, X\to N, (A\to N)\to B\Rightarrow C$\\
			$\Delta, (A\to B)\to X\Rightarrow C$&$\rightsquigarrow$&$\Delta, N\to X, (A\to B)\to N\Rightarrow C$\\
		\end{tabular}
	\end{center}

@@

	\begin{lemma}
		Iteratively applying the above transformations we can obtain from every sequent $S$ an equi-provable implicational sequent $S_\to$.
	\end{lemma}

		Check that for any formula that is not implicational there is an applicable rule.

		Check that the sequents on both sides of each rule are equi-provable.

		Termination can be shown by well-foundedness: First the number of formulas in the succedent is reduced to $1$, then the succedent is reduced to an atom. Then in each step the number of disjunctions, conjunctions or the depth of nested implications is reduced. 
	\end{proof}

>>>>

\usepackage{hyperref}% http://ctan.org/pkg/hyperref
\usepackage{cleveref}% http://ctan.org/pkg/cleveref

@@

\DeclareMathOperator*{\argmax}{arg\,max}

@@

	A sequent $A\Rightarrow B$ consists of two finite multi-sets of formulas $A, B$. We will be using the following rules for our classical calculus:\\
	\begin{center}
	\begin{tabular}{lll}
		\AxiomC{\hphantom{x}}
		\RightLabel{Ax}
		\UnaryInfC{$A,\Gamma\Rightarrow \Delta, A$}
		\DisplayProof&
		\AxiomC{\hphantom{x}}
		\RightLabel{L$\bot$}
		\UnaryInfC{$\bot,\Gamma\Rightarrow\Delta$}
		\DisplayProof&
		\\&&\\
		\AxiomC{$A, B,\Gamma\Rightarrow\Delta$}
		\RightLabel{L$\wedge$}
		\UnaryInfC{$A\wedge B, \Gamma\Rightarrow \Delta$}
		\DisplayProof&
		\AxiomC{$\Gamma\Rightarrow\Delta, A$}
		\AxiomC{$\Gamma\Rightarrow\Delta, B$}
		\RightLabel{R$\wedge$}
		\BinaryInfC{$\Gamma\Rightarrow \Delta, A\wedge B$}
		\DisplayProof&
		\\&&\\
		\AxiomC{$A, \Gamma\Rightarrow\Delta$}
		\AxiomC{$B, \Gamma\Rightarrow\Delta$}
		\RightLabel{L$\vee$}
		\BinaryInfC{$A\vee B, \Gamma\Rightarrow \Delta$}
		\DisplayProof&
		\AxiomC{$\Gamma\Rightarrow\Delta, A, B$}
		\RightLabel{R$\vee$}
		\UnaryInfC{$\Gamma\Rightarrow \Delta, A\vee B$}
		\DisplayProof&
		\\&&\\
		\AxiomC{$A\to B, \Gamma\Rightarrow\Delta, A$}
		\AxiomC{$B, \Gamma\Rightarrow\Delta$}
		\RightLabel{L$\to$}
		\BinaryInfC{$A\to B, \Gamma\Rightarrow \Delta$}
		\DisplayProof&
		\AxiomC{$A,\Gamma\Rightarrow\Delta, B$}
		\RightLabel{R$\to$}
		\UnaryInfC{$\Gamma\Rightarrow \Delta, A\to B$}
		\DisplayProof&
		\\&&\\
	\end{tabular}
	\end{center}

@@

	\begin{definition}
		We say that two sequents $S, S'$ are \textit{equiprovable} if $S$ is intuitionistically provable iff $S'$ is intuitionistically provable and write $$S\sim_p S'.$$
	\end{definition}

	Clearly equiprovability is an equivalence relation. Since every sequent is either provable or not the above notion is trivial in a sense. However we can often argue that two sequents are equiprovable even when we cannot determine their individual provability. This is in particular useful when discussing transformations of sequents that are to preserve intuitionistic provability.

	\begin{example}
		The sequents $\Rightarrow A\vee B$ and $A\to C, B\to C\Rightarrow C$ are equiprovable. Note that they are not equivalent in the typical sense however, due to the absence of $C$ in the former. 
	\end{example}

@@

		However there is clearly some connection between the two notions:

		\begin{lemma}\label{ep1}
			Let $A$ and $B$ be equivalent formulas, i.e. $A\Rightarrow B$ and $B\Rightarrow A$ are provable. Then \begin{enumerate}
				\item $A,\Gamma\Rightarrow\Delta \sim_p B, \Gamma\Rightarrow\Delta$
				\item $\Gamma\Rightarrow\Delta, A\sim_p\Gamma\Rightarrow\Delta, B$
			\end{enumerate}
		\end{lemma}

@@

	\begin{lemma}\label{ep2}
		Let $N$ be a atomic proposition that does not occur in $\Gamma$ or $\Delta$, $A, B$ arbitrary formulas that do not contain $N$, then:
		\begin{enumerate}
			\item $\Gamma\Rightarrow\Delta\sim_p\Gamma\Rightarrow\Delta, N$
			\item $\Gamma\Rightarrow\Delta, A\sim_p\Gamma, A\to N\Rightarrow \Delta, N$
			\item $\Gamma, A, B\Rightarrow\Delta\sim_p\Gamma, A\wedge B\Rightarrow\Delta$
			\item $\Gamma\Rightarrow A\vee B,\Delta\sim_p\Gamma\Rightarrow A, B,\Delta$
		\end{enumerate} 
	\end{lemma}

@@

		2. If $\Gamma\Rightarrow\Delta, A$ is provable we obtain $\Gamma, A\to N\Rightarrow \Delta, N$ through a simple application of cut:
		\begin{center}
			\AxiomC{$\vdots$}
			\noLine
			\UnaryInfC{$\Gamma\Rightarrow\Delta, A$}
			\AxiomC{\hphantom{X}}
			\RightLabel{Ax}
			\UnaryInfC{$A, A\to N\Rightarrow P, A$}
			\AxiomC{\hphantom{X}}
			\RightLabel{Ax}
			\UnaryInfC{$N, A\Rightarrow N$}
			\RightLabel{L$\to$}
			\BinaryInfC{$A, A\to N\Rightarrow N$}
			\RightLabel{Cut}
			\BinaryInfC{$\Gamma, A\to N\Rightarrow\Delta, N$}
			\DisplayProof
		\end{center}
		Suppose on the other hand we have a proof of $\Gamma, A\to N\Rightarrow \Delta, N$. Consider the proof obtained by replacing every occurrence of $N$ with $A$. By induction on the proof height it can be checked that this constitutes a valid proof of $\Gamma, A\to A\Rightarrow\Delta, N$. We obtain $\Gamma\Rightarrow\Delta, A$ through an application fo Cut:
		\begin{center}
			\AxiomC{\hphantom{X}}
			\RightLabel{Ax}
			\UnaryInfC{$A \Rightarrow A$}
			\RightLabel{R$\to$}
			\UnaryInfC{$\Rightarrow A\to A$}
			\AxiomC{$\vdots$}
			\noLine
			\UnaryInfC{$\Gamma, A\to A\Rightarrow\Delta, A$}
			\RightLabel{Cut}
			\BinaryInfC{$\Gamma\Rightarrow\Delta, A$}
			\DisplayProof
		\end{center}

		3. Going from $\Gamma, A, B\Rightarrow\Delta$ to $\Gamma, A\wedge B\Rightarrow\Delta$ is a single $L\wedge$ step. The other direction is a simple application of Cut:

		\begin{center}
			\AxiomC{\hphantom{X}}
			\RightLabel{Ax}
			\UnaryInfC{$A, B \Rightarrow A$}
			\AxiomC{\hphantom{X}}
			\RightLabel{Ax}
			\UnaryInfC{$A, B \Rightarrow B$}
			\RightLabel{R$\wedge$}
			\BinaryInfC{$A, B \Rightarrow A\wedge B$}
			\AxiomC{$\vdots$}
			\noLine
			\UnaryInfC{$\Gamma, A\wedge B\Rightarrow\Delta$}
			\RightLabel{Cut}
			\BinaryInfC{$\Gamma, A, B\Rightarrow\Delta$}
			\DisplayProof
		\end{center}

		4. Analogous to 3.
	\end{proof}

@@

	The first result is that for every sequent there is some equiprovable implicational sequent. This transformation is not strictly necessary for what follows, but it makes the arguments and procedure considerably simpler. It is reminiscent of the definitions of connectors in second order logic.

	Consider the following rules, where $X$ is not atomic, $N$ is a new atom:
	\begin{center}
		\begin{tabular}{lcl}
			$\Delta\Rightarrow \Gamma, A, B$&$\rightsquigarrow$&$\Delta\Rightarrow\Gamma, A\vee B$\\
			$\Delta\Rightarrow A\to B$&$\rightsquigarrow$&$\Delta, A\Rightarrow B$\\
			$\Delta\Rightarrow A\vee B$&$\rightsquigarrow$&$\Delta, (A\vee B)\to N\Rightarrow N$\\
			$\Delta\Rightarrow A\wedge B$&$\rightsquigarrow$&$\Delta,(A\wedge B)\to N\Rightarrow N$\\

			$\Delta, A\wedge B\Rightarrow C$&$\rightsquigarrow$&$\Delta, A, B\Rightarrow C$\\
			$\Delta, A\vee B\Rightarrow C$&$\rightsquigarrow$&$\Delta, (A\to C)\to (B\to C)\to N\Rightarrow  N$\\
			$\Delta, (A\wedge B)\to C\Rightarrow D$&$\rightsquigarrow$&$\Delta, A\to (B\to C)\Rightarrow D$\\
			$\Delta, (A\vee B)\to C\Rightarrow D$&$\rightsquigarrow$&$\Delta, A\to C, B\to C\Rightarrow D$\\
			$\Delta, A\to (B\vee C)\Rightarrow D$&$\rightsquigarrow$&$\Delta, A\to (B\to D)\to (C\to D)\to N\Rightarrow N$\\
			$\Delta, A\to (B\wedge C)\Rightarrow D$&$\rightsquigarrow$&$\Delta, A\to B, A\to C\Rightarrow D$\\
			$\Delta, X\to A\to B\Rightarrow C$&$\rightsquigarrow$&$\Delta, X\to N, N\to (A\to B)\Rightarrow C$\\
			$\Delta, A\to X\to B\Rightarrow C$&$\rightsquigarrow$&$\Delta, X\to N, A\to (N\to B)\Rightarrow C$\\
			$\Delta, A\to B\to X\Rightarrow C$&$\rightsquigarrow$&$\Delta, N\to X, A\to (B\to N)\Rightarrow C$\\
			$\Delta, (X\to A)\to B\Rightarrow C$&$\rightsquigarrow$&$\Delta, N\to X, (N\to A)\to B\Rightarrow C$\\
			$\Delta, (A\to X)\to B\Rightarrow C$&$\rightsquigarrow$&$\Delta, X\to N, (A\to N)\to B\Rightarrow C$\\
			$\Delta, (A\to B)\to X\Rightarrow C$&$\rightsquigarrow$&$\Delta, N\to X, (A\to B)\to N\Rightarrow C$\\
		\end{tabular}
	\end{center}

@@

	\begin{lemma}
		Iteratively applying the above transformations we can obtain from every sequent $S$ an equiprovable implicational sequent $S_\to$.
	\end{lemma}

	\begin{proof}[Proof]

		There are three things to check: 1. There is an applicable transformation for every non-implicational sequent. 2. The translation procedure terminates on every input. 3. The sequents involved in every transformation are equiprovable.

		1. Suppose $\Gamma\Rightarrow\Delta$ is not implicational. Then the succedent is not a single atom, or there is a formula in the antecedent that is not of the form $(C\to D)\to E, C\to (D\to E), C\to D$ or $D$ for some atoms $C, D, E$. In the former case one of transformations 1-4 is applicable. In the latter case one of the other rules is.

		2. Define the complexity $c(\varphi)$ of a formula as $c(A) = 0$ for atoms $A$, $c(A\circ B) = c(A) + c(B) + 1$ for $\circ\in\{\wedge,\vee,\to\}$. We can use $c$ to define a partial order $\succ$ on finite multi-sets of formulas, i.e. have $\emptyset\not \succ\Delta$ and $\Gamma\succ\emptyset$ for $\Gamma$ non-empty and for both $\Gamma,\Delta$ non-empty have\begin{align*}
			\Gamma \succ\Delta:\Longleftrightarrow&\max_\Gamma c > \max_\Delta c\text{ or }\\&\max_\Delta c =\max_\Gamma c\text{ and }\Gamma - \argmax_\Gamma c \succ\Delta - \argmax_\Delta c.
		\end{align*} $\succ$ is well-founded: Suppose towards contradiction that there is an infinite chain $$\Gamma_0^{(0)} \succ \Gamma_1^{(0)} \succ\dots\succ\Gamma_n^{(0)}\succ\dots$$Then every $\Gamma_i^{(0)}$ must be non-empty and $$\max_{\Gamma_0^{(0)}}c \geq \max_{\Gamma_1^{(0)}}c \geq\dots \geq \max_{\Gamma_n^{(0)}}c \geq\dots$$ and since $\mathbb{N}$ is well-ordered this chain must become stationary, i.e. there is $N_0$ such that $\max_{\Gamma_i^{(0)}}c = \max_{\Gamma_{N_0}^{(0)}}c$ for $i\geq N_0$ and $\max_{\Gamma_i^{(0)}}c>\max_{\Gamma_{N_0}^{(0)}}c$ for $i<N_0$. In particular for $i\geq N_0$ we have $\Gamma_i^{(0)}-\argmax_{\Gamma_i^{(0)}}c >_c \Gamma_{i+1}^{(0)}-\argmax_{\Gamma_{i+1}^{(0)}}c$, thus we obtain a second chain $$\Gamma_0^{(1)} \succ\Gamma_1^{(1)} \succ\dots\succ\Gamma_n^{(1)}\succ\dots$$ with $$\Gamma_i^{(1)} = \Gamma_{N+i}^{(0)} - \argmax_{\Gamma_{N+i}^{(0)}}c.$$We can repeat this construction to obtain an sequence $N_0, N_1,\dots$ Note that $$ \max_{\Gamma_{N_i}^{(i)}}c\geq \max_{\Gamma_{0}^{(i+1)}}c\geq \max_{\Gamma_{N_{i+1}}^{(i+1)}}c$$where the last is even a strict inequality if $N_{i+1} > 0$. Observe that $\Gamma_{N_0+\dots+N_i}^{(0)}$ by construction must contain at least $i+1$ formulas, i.e. $N_0,N_1,\dots$ cannot become constant 0. Consider a subsequence $N_{\alpha(0)},N_{\alpha(1)},\dots$ that does not contain a zero. Then we have $$\max_{\Gamma_{N_{\alpha(0)}}^{({\alpha(0)})}}c > \max_{\Gamma_{N_{\alpha(1)}}^{({\alpha(1)})}}c>\dots >\max_{\Gamma_{N_{\alpha(n)}}^{({\alpha(n)})}}c>\dots$$ which is an infinite decreasing chain, but $\mathbb{N}$ is a well-order. $\lightning$

	 	We can now use $\succ$ to define a partial order $\gg$ on the set of sequents as follows: $$\Gamma\Rightarrow\Delta \gg \Gamma'\Rightarrow\Delta'$$ iff
		\begin{enumerate}
			\item $|\Delta| > |\Delta'|$ 
			\item not 1 and $\Delta \succ \Delta'$
			\item not 1, 2 and the number of occurrences of $\wedge$ in $\Gamma$ is higher than in $\Gamma' $.
			\item not 1 - 3 and the number of occurrences of $\vee$ in $\Gamma$ is higher than in $\Gamma' $.
			\item not 1 - 4 and $\Gamma \succ\Gamma' $.
		\end{enumerate}
		Since each of 1 - 5 is well-founded, $\gg$ is well-founded, i.e. for each set we can obtain a minimal element by considering first a minimal subset with regard to 1, then to 2 and so on.

		Now every transformation is reductive with regard to $\gg$. Therefore iteratively applying transformations creates a descending chain. Since every descending chain in a well-founded order is finite this proves termination.

		3. For most cases equiprovability follows from Lemmas \ref{ep1} and \ref{ep2}, e.g. consider $$\Delta, A\to (B\wedge C)\Rightarrow D\indent\rightsquigarrow\indent\Delta, A\to B, A\to C\Rightarrow D$$Then we have
		\begin{align*}
			\Delta, A\to (B\wedge C)\Rightarrow D&\overset{\ref{ep1}.1}{\sim_p}\Delta, (A\to B)\wedge (A\to C)\Rightarrow D\\&\overset{\ref{ep2}.3}{\sim_p}\Delta, A\to B, A\to C\Rightarrow D
		\end{align*}

		Some cases require simple applications of cut using intuitionistic tautologies, e.g. for
		$$\Delta, A\vee B\Rightarrow C\indent\rightsquigarrow\indent\Delta, (A\to C)\to (B\to C)\to N\Rightarrow  N$$
		we have 
		\begin{align*}
			\Delta, A\vee B\Rightarrow C&\overset{\star}{\sim_p}\Delta\Rightarrow (A\vee B)\to C\\&\overset{\ref{ep2}.3}{\sim_p}\Delta\Rightarrow (A\to C)\wedge(B\to C)\\&\overset{\ref{ep2}.2}{\sim_p} \Delta, ((A\to C)\wedge (B\to C))\to N\Rightarrow N\\&\overset{\ref{ep1}.1}{\sim_p}\Delta, (A\to C)\to(B\to C)\to N\Rightarrow N.
		\end{align*}
		Where one direction of $\star$ is trivial and the other is obtained via cut with $$A\vee B, (A\vee B)\to C\Rightarrow C$$

		For the last 6 rules one direction in obtained via cut and the other one by substituting all occurrences of $N$ in a proof with $X$.
	\end{proof}
